---
title: "Quant DO-MS"
author: ""
date: "11/29/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source('QuantQC_functions.R')


# Read in the path to maxQuant file
#data_path <- '/Users/andrewleduc/Library/CloudStorage/GoogleDrive-leduc.an@husky.neu.edu/.shortcut-targets-by-id/1uQ4exoKlaZAGnOG1iCJPzYN3ooYYZB7g/MS/cellenONE/Trachea_3_7_pSCoPE/Trachea_pSCoPE_prep1_part2/evidence.txt'

#data_path2 <- '/Users/andrewleduc/Library/CloudStorage/GoogleDrive-leduc.an@husky.neu.edu/.shortcut-targets-by-id/1uQ4exoKlaZAGnOG1iCJPzYN3ooYYZB7g/MS/cellenONE/Trachea_3_7_pSCoPE/Trachea_3_7_pSCoPE_part1/evidence.txt'


data_path <- '/Users/andrewleduc/Library/CloudStorage/GoogleDrive-leduc.an@husky.neu.edu/.shortcut-targets-by-id/1uQ4exoKlaZAGnOG1iCJPzYN3ooYYZB7g/MS/cellenONE/Trachea_3_6_pSCoPE/Trachea_3_6_pSCoPE_part1/evidence.txt'



## Read in meta data (Just links raw file to inject well)
linker <- read.csv('/Users/andrewleduc/Library/CloudStorage/GoogleDrive-leduc.an@husky.neu.edu/.shortcut-targets-by-id/1uQ4exoKlaZAGnOG1iCJPzYN3ooYYZB7g/MS/cellenONE/Trachea_3_6_pSCoPE/linker_age.csv')



### Read in cell isolation files from CellenONE and assign cell type


one <-read.delim("/Users/andrewleduc/Library/CloudStorage/GoogleDrive-leduc.an@husky.neu.edu/.shortcut-targets-by-id/1uQ4exoKlaZAGnOG1iCJPzYN3ooYYZB7g/MS/cellenONE/Trachea_3_6_pSCoPE/Young_isolated.xls")
one$condition <- 'young'


two <- read.delim("/Users/andrewleduc/Library/CloudStorage/GoogleDrive-leduc.an@husky.neu.edu/.shortcut-targets-by-id/1uQ4exoKlaZAGnOG1iCJPzYN3ooYYZB7g/MS/cellenONE/Trachea_3_6_pSCoPE/old_isolated.xls")
two$condition <- 'old'


all_cells <- rbind(one,two)


# Set parameters

PIF <- .75
PEP <- .04

CV_thresh <- .43

#species <- c('Mouse') #Mouse, Human


```

# Results {.tabset}

```{r Pre_proccess, include=FALSE}


#Read and format raw data, link MS runs to injection wells
Trachea_3_7_prot <- MQ_to_nPOP(data_path,linker,PIF,PEP)


# Normalize single cell runs to reference channel, 
# filter out data points over twice reference
# Turn long table format to Peptide X Cell matrix
Trachea_3_7_prot <- TMT_Reference_channel_norm(Trachea_3_7_prot)



## mapping cellenONE meta data to raw data
Trachea_3_7_prot <- link_cellenONE_Raw(Trachea_3_7_prot,all_cells)


```


## LC/MS Performance Statistics

### IDs and intensities
```{r, echo=FALSE,fig.width = 10, fig.height = 4.5}

# Calculate statistics to look for decrease in LC/MS performance
# over time
Trachea_3_7_prot <- Calculate_run_order_statistics(Trachea_3_7_prot)


# Plot MS Intensity Statistics
PlotIntensityDrift(Trachea_3_7_prot)


```


### Retention time drift
```{r, echo=FALSE,fig.width = 10, fig.height = 4.5}

#Plot Retention time Statistics
PlotRTDrift(Trachea_3_7_prot)


```


## Single Cell Statistics


### Negative control vs real cell statistics
```{r negative_ctrl ,echo=FALSE, fig.width = 11, fig.height = 4.5,warning=FALSE,message=FALSE}

PlotNegCtrl(Trachea_3_7_prot,.43)

Trachea_3_7_prot <- FilterFailedCells(Trachea_3_7_prot)

```

### Carrier Statistics

```{r carrier, echo=FALSE, fig.height=4.5, fig.width=11, warning=FALSE,message=FALSE}

#Compute the size relative to the carrier of each cell
PlotSCtoCarrierRatio(Trachea_3_7_prot)

```


### Protein and peptide numbers per cell

```{r Collapse and count, echo=FALSE,fig.width = 11, fig.height = 4.5,warning=FALSE,message=FALSE}

Trachea_3_7_prot <- CollapseToProtein(Trachea_3_7_prot,1)

PlotProtAndPep(Trachea_3_7_prot)

```


### Data completness based on filtering (uses intersected subset of proteins from channel qval filtered data)


```{r completness_Statistics, echo=FALSE,fig.width = 11.5, fig.height = 4.5,warning=FALSE,message=FALSE}

PlotDataComplete(Trachea_3_7_prot)

```

### Correlations between peptides mapping to protein compared to variance

```{r correlations_pep to prot, echo=FALSE,fig.width = 7, fig.height = 4.5}

#Compute correlations between peptides mapping to same protein

pep_cor <- Pep_cor(Ref_norm_data_filtered,protein_mat,F,cellenONE_meta)


pep_cor_plot <- ggplot(pep_cor, aes(y = Cor, x = FC)) + geom_boxplot() + xlab('Mean abs(protein fold change)') +ylab('Correlation between peptides mapping to a protein')+stat_summary(fun.data=f, geom="text", vjust=-0.5, col="blue")

pep_cor_plot


```

### PCA and batch effects checks
```{r PCA, echo=FALSE, fig.height=4.5, fig.width=11, message=FALSE, warning=FALSE, include=TRUE}

# Impute data (needed for batch correction)
protein_mat_imputed <- KNN_impute(protein_mat,3)









# Regular PCA by sample
ggplot(scx,aes(x = PC1,y = PC2, color = sample)) + geom_point(size = 5) + theme_classic()+ ggtitle('Cell Type') +
  xlab(paste0('PC1 ',round(percent_var[1]),'%'))+ylab(paste0('PC2 ',round(percent_var[2]),'%'))



# PCA by labels to check label bias
lab_plot <- ggplot(scx,aes(x = PC1,y = PC2, color = label)) + geom_point(size = 5) + theme_classic() + ggtitle('Label Bias')

# PCA by intensity to check size bias
intense_plot <-  ggplot(scx,aes(x = PC1,y = PC2, color = prot_total)) + geom_point(size = 5) + theme_classic()+ ggtitle('Cell Intensity')+
  scale_color_gradient2(midpoint = median(scx$prot_total,na.rm = T), low = "blue", mid = "white",high = "red",name = '')

lab_plot 
intense_plot

Run_order_batch <- ggplot(scx,aes(x = PC1,y = PC2, color = Order)) + geom_point(size = 5) + theme_classic()+ ggtitle('Order sets were ran in')+
  scale_color_gradient2(midpoint = median(scx$Order,na.rm = T), low = "blue", mid = "white",high = "red",name = '')

Run_order_batch

# Distribution of cell sizes by diameter
ggplot(scx,aes(x = sample,y = diameter)) + geom_boxplot() + 
  theme_classic()+ ggtitle('Size by sample')+ylab('Diameter (micron)')+xlab('')+
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))


# Examine cell volume to total protein relationship, should be linear correlation

scx$vol <- ((scx$diameter/2)^3)

ggplot(scx,aes(x = vol,y = prot_total,color=sample)) + geom_point(size = 5) + theme_classic()+ ggtitle(paste0('Correlation = ',round(cor(scx$prot,scx$vol,use = 'pairwise.complete.obs', method = 'pearson'),digits = 3)))+ylab('Total (Summed) Cell Intensity')+xlab('Cell Volume (diameter/2)^3')


ggplot(scx,aes(x = vol,y = prot_total,color=Order)) + geom_point(size = 5) + theme_classic()+ ggtitle('Colored by order sets were run')+ylab('Total (Summed) Cell Intensity')+xlab('Cell Volume (diameter/2)^3')+
 scale_color_gradient2(midpoint = median(scx$Order,na.rm = T), low = "blue", mid = "white",high = "red",name = '')




```



### Additional dimensionality reduction and clustering


```{r UMAP, echo=FALSE, fig.height=4.5, fig.width=11, message=FALSE, warning=FALSE, include=TRUE}



 
rownames(sc.batch_cor) <- rownames(protein_mat_imputed)
#sc.batch_cor <- sc.batch_cor[rownames(sc.batch_cor)%in% convert_mouse_have$Entry, ]
#sc.batch_cor <- sc.batch_cor[convert_mouse_have$Entry,]
#rownames(sc.batch_cor) <- convert_mouse_have$Gene.Names..synonym.


protein_Data <- XXX # Normalized prote data I sent

prot_umap <- CreateSeuratObject(counts = protein_Data, project = "prot_mat")
prot_umap <- NormalizeData(prot_umap, normalization.method = "LogNormalize", scale.factor = 10000)
prot_umap@assays$RNA@scale.data <- sc.batch_cor

all.genes <- rownames(protein_Data)
prot_umap <- ScaleData(prot_umap, features = protein_Data)
prot_umap@assays$RNA@scale.data <- protein_Data
prot_umap <- RunPCA(prot_umap, features = all.genes)

DimHeatmap(prot_umap, dims = 1:6, cells = 200, balanced = TRUE)
prot_umap <- FindNeighbors(prot_umap, dims = 1:6)
prot_umap <- FindClusters(prot_umap, resolution = 1.5)
prot_umap <- RunUMAP(prot_umap, dims = 1:6)
DimPlot(prot_umap, reduction = "umap") + ggtitle('UMAP colored by Louvain cluster')


um_plot <- as.data.frame(prot_umap@reductions[["umap"]]@cell.embeddings)
PCA_plot <- as.data.frame(prot_umap@reductions[["pca"]]@cell.embeddings)

um_plot$Celltype <- scx$sample
um_plot$cluster <- prot_umap@meta.data[["RNA_snn_res.1.5"]]
um_plot$lab <- scx$label
um_plot$Order <- scx$Order

PCA_plot$Celltype <- scx$sample
PCA_plot$lab <- scx$label
PCA_plot$Order <- scx$Order

ggplot(um_plot, aes(x = UMAP_1,y = UMAP_2, color = cluster)) + geom_point(size = 3) + theme_classic()+
  ggtitle('UMAP colored by Sample')




#um_plot$Mark1 <- sc.batch_cor_noimp['O88312',]
#ggplot(um_plot, aes(x = UMAP_1,y = UMAP_2, color = Mark1)) + geom_point(size = 3) + theme_classic()+
#  scale_color_gradient2(midpoint = 0, low = "blue", mid = "white",high = "red",name #= '')

# pvals <- c()
# prot_names <- c()
# for(i in 1:nrow(sc.batch_cor_noimp)){
#   C1 <- sc.batch_cor_noimp[i,prot_umap@meta.data[["seurat_clusters"]]%in% c(0,3)]
#   C2 <- sc.batch_cor_noimp[i,prot_umap@meta.data[["seurat_clusters"]]%in% c(1,2)]
#   
#   if(sum(is.na(C1)==F) > 9 & sum(is.na(C2)==F) > 9 ){
#     pval <- t.test(C1,C2)$p.value
#     
#     pvals <- c(pvals,pval)
#     prot_names <- c(prot_names,rownames(sc.batch_cor_noimp)[i])
#     
#     
#   }
#   
# }
# 
# df <- as.data.frame(pvals)
# df$prot <- prot_names

```




### Advanced Peptide mapping to a Protein Correlations

```{r r Advanced Pep to prot Correlations, echo=FALSE, fig.height=4.5, fig.width=11, message=FALSE, warning=FALSE, include=TRUE}

# 
# cells_filt_adj <- as.matrix(Ref_norm_data_filtered[,3:ncol(Ref_norm_data_filtered)])
# 
# cells_filt_adj_imp <- KNN_impute(cells_filt_adj,3)
# cells_filt_adj_imp <- limma::removeBatchEffect(cells_filt_adj_imp,batch = batch_label$injectWell, batch2 = batch_label$label)
# cells_filt_adj_imp[is.na(cells_filt_adj)==T] <- NA
# 
# 
# 
# 
# clust_ids <- as.numeric(as.character(um_plot$assign))
# #clust_ids <- as.numeric(as.character(prot_umap@meta.data[["RNA_snn_res.1.5"]]))
# 
# new_mat <- matrix(data = NA,nrow = nrow(Ref_norm_data_filtered),ncol = length(unique(clust_ids)))
# 
# 
# 
# count= 0
# for(i in 1:length(unique(clust_ids))){
#   new_mat[,i] <- rowMeans(cells_filt_adj_imp[,clust_ids == count],na.rm = T)
# 
#   count = count + 1
# }
# new_mat[is.nan(new_mat)] <- NA
# 
# new_mat <- cbind(Ref_norm_data_filtered[,1:2],new_mat)
# 
# pep_cor <- Pep_cor(new_mat,protein_mat)
# 
# 
# 
# pep_cor_plot <- ggplot(pep_cor, aes(y = Cor, x = FC)) + geom_boxplot() + xlab('Mean abs(protein fold change)') +ylab('Correlation peptides from a protein, cluster level')+stat_summary(fun.data=f, geom="text", vjust=-0.5, col="blue") + ggtitle('Correlations averaged at the cluster level')
# 
# pep_cor_plot


```




```{r Plotting_Stats, echo=FALSE,fig.width = 11, fig.height = 4.5}


# prot_info <- prot_info %>% filter(is.na(vars) == F)
# Var_norm <- ggplot(prot_info,aes(x = log10(sums),y = log10(vars))) + geom_point() + 
#   theme_bw()  + ggtitle(paste0('No normalization, ',"Cor = ", round(cor(log10(prot_info$sums),log10(prot_info$vars),use = 'pairwise.complete.obs'),digits = 3))) + xlab('Mean Protein Intensity') + ylab('Protein Variance')+ theme(axis.title.x = element_text(size = 20),axis.title.y = element_text(size = 20),axis.text.x = element_text(size = 15),axis.text.y = element_text(size = 15))
# 
# Var_no_norm <-ggplot(prot_info,aes(x = log10(sums),y = log10(vars_no_norm))) + geom_point() + 
#   theme_bw() + ggtitle(paste0('After normalization, ' ,"Cor = ", round(cor(log10(prot_info$sums),log10(prot_info$vars_no_norm),use = 'pairwise.complete.obs'),digits = 3)))+ xlab('Mean Protein Intensity') + ylab('Protein Variance')+ theme(axis.title.x = element_text(size = 20),axis.title.y = element_text(size = 20),axis.text.x = element_text(size = 15),axis.text.y = element_text(size = 15))
# 
# Var_no_norm + Var_norm


```



```{r, echo=FALSE,fig.width = 11, fig.height = 4.5}

#Saads plots
# Norm_quality_pre_cell <-ggplot(cell_info,aes(x = log10(sum_no_norm),y = complete)) + geom_point() + ylab('% Completeness of proteins in cell')+ xlab('log10(summed Intensity of cell)')+ theme_bw()+ ggtitle(paste0('No normalization, Cor = ', round(cor(log10(cell_info$sum_no_norm),cell_info$complete,use = 'pairwise.complete.obs'),digits = 3))) +theme(axis.title.x = element_text(size = 15),axis.title.y = element_text(size = 15),axis.text.x = element_text(size = 10),axis.text.y = element_text(size = 10)) 
# 
# Norm_quality_post_cell_saad <-ggplot(cell_info,aes(x = (sum_norm),y = complete)) + geom_point() + theme_bw()+ ggtitle(paste0('New normalization, Cor = ', round(cor(cell_info$sum_norm,cell_info$complete,use = 'pairwise.complete.obs'),digits = 3))) + ylab('% Completeness of proteins in cell')+ xlab('log10(summed Intensity of cell)')+ theme(axis.title.x = element_text(size = 15),axis.title.y = element_text(size = 15),axis.text.x = element_text(size = 10),axis.text.y = element_text(size = 10))
# 
# Norm_quality_post_cell_old <-ggplot(cell_info,aes(x = (sum_norm_old),y = complete)) + geom_point() + ggtitle(paste0('Old normalization, Cor = ', round(cor(cell_info$sum_norm_old,cell_info$complete,use = 'pairwise.complete.obs'),digits = 3)))+ theme_bw() + theme(axis.title.x = element_text(size = 15),axis.title.y = element_text(size = 15),axis.text.x = element_text(size = 10),axis.text.y = element_text(size = 10))
# 
# Norm_quality_pre_cell + Norm_quality_post_cell_old + Norm_quality_post_cell_saad
```



```{r, echo=FALSE,fig.width = 11, fig.height = 4.5}

# Norm_quality_pre_prot <-ggplot(prot_info,aes(x = log10(sums),y = complete)) + geom_point() + 
#   theme_bw() + ggtitle('No normalization')+ xlab('Mean Protein Intensity') + ylab('# Complete')+ theme(axis.title.x = element_text(size = 15),axis.title.y = element_text(size = 15),axis.text.x = element_text(size = 10),axis.text.y = element_text(size = 10))
# 
# Norm_quality_post_prot <-ggplot(prot_info,aes(x = (sum_norm),y = complete)) + geom_point() + 
#   theme_bw()+ggtitle('With normalization')+ xlab('Mean Protein Intensity') + ylab('# Complete')+ theme(axis.title.x = element_text(size = 15),axis.title.y = element_text(size = 15),axis.text.x = element_text(size = 10),axis.text.y = element_text(size = 10))
# 
# 
# Norm_quality_post_prot_var <-ggplot(prot_info,aes(x = log2(vars),y = complete)) + geom_point() + theme_bw() +ggtitle('With normalization')+ xlab('Protein Variance') + ylab('# Complete')+ theme(axis.title.x = element_text(size = 15),axis.title.y = element_text(size = 15),axis.text.x = element_text(size = 10),axis.text.y = element_text(size = 10))
# 
#   
#   #annotate("text", x=.75, y=3,size=12, label= paste0("Cor = ", round(cor(log10(prot_info$sums),log10(prot_info$vars_no_norm),use = 'pairwise.complete.obs'),digits = 3)))
# 
# Norm_quality_pre_prot + Norm_quality_post_prot + Norm_quality_post_prot_var

```

## CellenONE_info

### Cell type positions across slide
```{r, echo=FALSE,fig.width = 15, fig.height = 10}

# print the cell IDs overlayed on the positions of the slide
ggplot(Trachea_3_7_prot@cellenONE.meta) + 
  geom_point(aes(x = dropXPos,y = dropYPos,color = sample)) + 
  geom_text(aes(x = pickupXPos_numb,y = pickupYPos_numb,label = injectWell,size = 5),hjust= .5, vjust=-.6) + 
  facet_wrap(~field,ncol = 4)+
  scale_y_reverse() 




```

### Label position across slide
```{r, echo=FALSE,fig.width = 13, fig.height = 10}
# print the mTRAQ labels overlayed on the positions of the slide
ggplot(Trachea_3_7_prot@cellenONE.meta, aes(x = dropXPos,y = dropYPos,color = label)) + 
   geom_point() +scale_y_reverse()+ facet_wrap(~field,ncol = 4)

```
