---
title: "Quant DO-MS"
author: ""
date: "11/29/2022"
output: html_document
params:
  data_path: "data/file1.csv"
  linker_path: "data/file2.csv"
  isolation: "data/file3.csv"
  prot_vis_umap: NULL

  
---

```{r setup, include=FALSE}


# Set parameters

PIF <- .75
PEP <- .04

CV_thresh <- .43

#species <- c('Mouse') #Mouse, Human


```

# Results {.tabset}

```{r Pre_proccess, include=FALSE}


#Read and format raw data, link MS runs to injection wells
nPOP_obj_Exp <- MQ_to_nPOP(params$data_path,params$linker_path,PIF,PEP)


# Normalize single cell runs to reference channel, 
# filter out data points over twice reference
# Turn long table format to Peptide X Cell matrix
nPOP_obj_Exp <- TMT_Reference_channel_norm(nPOP_obj_Exp)



## mapping cellenONE meta data to raw data
nPOP_obj_Exp <- link_cellenONE_Raw(nPOP_obj_Exp,params$isolation)


```


## LC/MS Performance Statistics

### IDs and intensities
```{r, echo=FALSE,fig.width = 10, fig.height = 12}

# Calculate statistics to look for decrease in LC/MS performance
# over time
nPOP_obj_Exp <- Calculate_run_order_statistics(nPOP_obj_Exp)


# Plot MS Intensity Statistics
PlotIntensityDrift(nPOP_obj_Exp)


```


### Retention time drift
```{r, echo=FALSE,fig.width = 10, fig.height = 8}

#Plot Retention time Statistics
PlotRTDrift(nPOP_obj_Exp)


```


## Single Cell Statistics


### Negative control vs real cell statistics
```{r negative_ctrl ,echo=FALSE, fig.width = 11, fig.height = 4.5,warning=FALSE,message=FALSE}

nPOP_obj_Exp <- EvaluateNegativeControls(nPOP_obj_Exp,.43)


PlotNegCtrl(nPOP_obj_Exp,.43)


nPOP_obj_Exp <- FilterBadCells(nPOP_obj_Exp,.43)

```

### Carrier Statistics

```{r carrier, echo=FALSE, fig.height=4.5, fig.width=11, warning=FALSE,message=FALSE}



#Compute the size relative to the carrier of each cell
PlotSCtoCarrierRatio(nPOP_obj_Exp)

```

### Ms1 vs Ms2 Peptide correlations

```{r carrier, echo=FALSE, fig.height=4.5, fig.width=11, warning=FALSE,message=FALSE}


PlotMS1vMS2(Trachea_3_7_prot)

```


### Protein and peptide numbers per cell

```{r Collapse and count, echo=FALSE,fig.width = 11, fig.height = 4.5,warning=FALSE,message=FALSE}

nPOP_obj_Exp <- CollapseToProtein(nPOP_obj_Exp,1)

PlotProtAndPep(nPOP_obj_Exp)

```


### Data completness based on filtering (uses intersected subset of proteins from channel qval filtered data)


```{r completness_Statistics, echo=FALSE,fig.width = 11.5, fig.height = 4.5,warning=FALSE,message=FALSE}

PlotDataComplete(nPOP_obj_Exp)

```

### Correlations between peptides mapping to protein compared to variance

```{r correlations_pep to prot, echo=FALSE,fig.width = 7, fig.height = 4.5}

#Compute correlations between peptides mapping to same protein

nPOP_obj_Exp <- SharedPeptideCor(nPOP_obj_Exp)

PlotPepCor(nPOP_obj_Exp)


```

### PCA and batch effects checks
```{r PCA, echo=FALSE, fig.height=4.5, fig.width=11, message=FALSE, warning=FALSE, include=TRUE}


# Impute, still old imputation very slow, soon will update with C++ implementation
nPOP_obj_Exp <- KNN_impute(nPOP_obj_Exp)


# currently does label and LC/MS run but options comming soon
nPOP_obj_Exp <- BatchCorrect(nPOP_obj_Exp)



nPOP_obj_Exp <- ComputePCA(nPOP_obj_Exp)

# plot PCA options are "Run order" "Total protein" "Condition" "Label"
PlotPCA(nPOP_obj_Exp, by = "Condition")


PlotPCA(nPOP_obj_Exp, by = "Label")


PlotPCA(nPOP_obj_Exp, by = "Run order")



```



### Additional dimensionality reduction and clustering


```{r UMAP, echo=FALSE, fig.height=4.5, fig.width=11, message=FALSE, warning=FALSE, include=TRUE}

nPOP_obj_Exp <- ComputeUMAP(nPOP_obj_Exp)

# plots by cluster
PlotUMAP(nPOP_obj_Exp)




```


### Viz proteins of interest
```{r Plotting_Stats, warning = FALSE, message = FALSE,echo=FALSE,fig.width = 11, fig.height = 4.5}


fig1 <- ProteinClustConsistency(nPOP_obj_Exp, prot = params$prot_vis_umap, type = 'line')
  

fig1

```

### Advanced Peptide mapping to a Protein Correlations

```{r r Advanced Pep to prot Correlations, echo=FALSE, fig.height=4.5, fig.width=11, message=FALSE, warning=FALSE, include=TRUE}


```



## CellenONE_info

### Cell type positions across slide
```{r, echo=FALSE,fig.width = 15, fig.height = 10}

# print the cell IDs overlayed on the positions of the slide

PlotSlideLayout_celltype(nPOP_obj_Exp)



```

### Label position across slide
```{r, echo=FALSE,fig.width = 13, fig.height = 10}

# print the cell IDs overlayed on the positions of the label
PlotSlideLayout_label(nPOP_obj_Exp)


```
